Future Parameter constraints from weak lensing CMB and Galaxy Lensing Power- and Bi-spectra

### 2025-06-18 paper/paper_draft.tex
added acknowledgment
added abstract

\emailAdd{jonasfrugte@gmail.com}
\emailAdd{p.d.meerburg@rug.nl}
\abstract{}

@@

>>>>

\emailAdd{jonasfrugte@gmail.com}
\emailAdd{p.d.meerburg@rug.nl}
\abstract{
Upcoming Stage 4 surveys from the Simons Observatory, LSST, and Euclid are poised to measure weak gravitational lensing of the Cosmic Microwave Background (CMB) and galaxies with unprecedented precision. While the power spectrum is the standard statistic used to analyze weak lensing data, non-Gaussianity induced by the non-linear growth of structure imparts significant cosmological information into higher-order statistics.

In this paper, we perform a Fisher matrix analysis to forecast the ability of future surveys to constrain cosmological parameters using the weak lensing power spectrum and bispectrum from both CMB and galaxy probes, as well as their cross-correlations. We analyze an eight-parameter cosmological model ($\Lambda$CDM + $\sum m_{\nu}$ + $w_{0}$) and assess parameter constraints for Stage 4 survey specifications.

Our analysis shows that, in the absence of systematics, both the CMB and galaxy lensing bispectra will be detectable at high signal-to-noise. We consider parameter constraints with a ``weak'' prior and with a stronger prior from the primary CMB anisotropies. In the weak prior case the lensing bispectrum substantially improves parameter constraints over using the lensing power spectrum alone, primarily by breaking key parameter degeneracies. For the stronger prior we see less significant gains, particularly for the CMB lensing bispectrum. We further find that for stage 4 surveys that are able to measure lensing on small scales where non-linear structure formation becomes important, the constraining power of the lensing bispectrum can become comparable to that of the lensing power spectrum. Furthermore, we demonstrate that, both for our weak and strong priors, the synergy between CMB and galaxy lensing is substantial. Combining the two probes, particularly with both their lensing power- and bi-spectra, leads to significantly tighter constraints than achievable with either probe individually, with the greatest improvement seen for the sum of neutrino masses.
}

@@

\section*{Acknowledgements}
We would like to thank prof. Toshiya Namikawa for providing extra details on his paper on the CMB lensing bispectrum \cite{Namikawa_2016}, which, in particular, allowed us to cross check (intermediate) results.
